---
title: "Fibroblast Trajectory"
output: html_document
header-includes:
  - \usepackage{subcaption}
  - \usepackage{float}
---
#### We conducted a trajectory analysis of all fibroblast cells in the integrated data.
#### There are two parts in this markdown file: 1) reclustering of all fibroblast cells; 2) trajectory analysis of fibroblast cells.
#### The integration was performed by rPCA in Seurat package. Louvain clustering was applied in the integrated data. Cell types and stages were showed in the integrated Umap (Figure 1.1-1.3).
#### The trajectory analysis was performed by slingshot based on clusters in the integrated data (Figure 2.1).
```{r, eval=F, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# Space separator
```

#### 1. Fibroblast in integrated data
```{r, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# fibroblast was taken as global cluster #2 (reso=.1)
fib_cl<- rownames(seu_clu)[seu_clu[,1]==2]
fib_mx<- lapply(mg_mx, function(x){ x[ ,colnames(x) %in% fib_cl ] })
# too few cells at batch-2 of our dataset, merged with batch-1
fib_mx[[1]]<- cbind( fib_mx[[1]], fib_mx[[2]])
fib_mx<- fib_mx[-2]
# divide Etc's dataset into stages (observe batch effect between stages if not)
fib_mx<-c(fib_mx[1:2], by( t(fib_mx[[3]]), all_meta[ colnames(fib_mx[[3]]), 'stage'], t)[c(2,3,1)] )

seu<- lapply( fib_mx, function(x){ CreateSeuratObject( x, min.cells = 0, min.features = 0 )})
seu <- mapply(function(x, y) {
    if(y) x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 1000)
}, x=seu, y=rep(T,5) )
length(features <- SelectIntegrationFeatures(object.list = seu,nfeatures=1000))
length( features<- setdiff( features, vis_rmg) )
seu <- lapply(X = seu, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})
anchor <- FindIntegrationAnchors(object.list = seu, anchor.features = features, reduction = "rpca", dims=1:50, reference=3)
seu_comb <- IntegrateData(anchorset = anchor)
DefaultAssay(seu_comb) <- "integrated"
# Run the standard workflow for visualization and clustering
seu_comb <- ScaleData(seu_comb, verbose = FALSE)
seu_comb <- RunPCA(seu_comb, npcs = 50, verbose = FALSE, features=features )
seu_comb <- RunUMAP(seu_comb, reduction = "pca", dims = 1:30)
fib_umap<-seu_comb@reductions$umap@cell.embeddings
fib_pcac<- seu_comb@reductions$pca@cell.embeddings
fib_hvg<- features
```

```{r, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# plot by dataset, cell type
rand_ind<- sample.int( nrow(fib_umap) )
tp3_col2<- tp3_col # change colors for fibroblast
tp3_col2[c(15:17)]<- col6[c(2:4)]
#pdf(paste('../etchevers/result/','fib/','gloclu2_stage.pdf',sep=''),width=7,height=7) # by all stages
par(cex=2,las=1,mar=rep(.5,4),lwd=3)
plot(fib_umap[rand_ind,1],fib_umap[rand_ind,2],pch=16,xlab='umap-1',ylab='umap-2',frame=F,col=st_col[as.character(all_meta[rownames(fib_umap),4]) ][rand_ind],cex=0.2, xaxt='n',yaxt='n')
legend( 'bottomright', col=st_col, legend=c(names(st_col)[1:3],'6.5-7w',names(st_col)[5:7]), xpd=NA, bty='n', pch=16, cex=.8)
#dev.off()
```
##### Figure 1.1: stages in the integrated data of fibroblast.
######## .

```{r, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# only plot fibroblast
tp3_col2[! names(tp3_col2) %in% c('Fb1','Fb2','Fb3','Fb4-VIC','SMC') ]<-NA
#pdf(paste('../etchevers/result/','fib/','gloclu2_dataset_fib.pdf',sep=''),width=7,height=7)
par(cex=2,las=1,mar=rep(.5,4),lwd=3)
plot(fib_umap[rand_ind,1],fib_umap[rand_ind,2],pch=16,xlab='umap-1',ylab='umap-2',frame=F,col=tp3_col2[as.character(all_meta[rownames(fib_umap),3]) ][rand_ind],cex=0.2,xaxt='n',yaxt='n',main='')
legend( 'bottomright', col=tp3_col2[c(14:17,21)], legend=names(tp3_col2)[c(14:17,21)], xpd=NA, bty='n', pch=16, cex=.8)
#dev.off()
```
##### Figure 1.2: cell types from this study in the integrated data of fibroblast.
######## .

```{r, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# clustering
seu_comb<- FindNeighbors(seu_comb, dims = 1:30)
fib_clus <- rep(0, nrow(fib_umap))
resos<- seq(.1,1,.1)
for( reso in resos){
  seu_comb<-FindClusters(seu_comb, resolution =reso, dims.use =1:30 )
  one_res<- as.numeric(seu_comb@active.ident)
  names(one_res)<- names(seu_comb@active.ident)
  fib_clus<- cbind( fib_clus, one_res)  
}
fib_clus<- fib_clus[,-1]
colnames(fib_clus)<- resos
rownames(fib_clus)<- rownames(fib_umap)
```

```{r, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# plot clusters
#pdf(paste('../etchevers/result/','fib/','gloclu2_seuclu.pdf',sep=''), width=8,height=12)
par(cex=2,las=1,mar=c(5,4,1,1),lwd=3, pch=16 )
i<-5
plot_cen<- sapply( sort(unique(fib_clus[,i])), function(x){
  cl<- rownames(fib_clus)[fib_clus[,i]==x]
  res<-c(median(fib_umap[cl,1]), median(fib_umap[cl,2]))
  return(res)
})
plot(fib_umap[rand_ind,1],fib_umap[rand_ind,2],pch=16,xlab='umap-1',ylab='umap-2',frame=F,col=col6[-1][fib_clus[,i]][rand_ind],cex=0.2,main=max(fib_clus[,i]))
for(i in 1:ncol(plot_cen)) text( plot_cen[1,i], plot_cen[2,i], label=i, xpd=NA)
#dev.off()
```
##### Figure 1.3: clustering in the integrated data of fibroblast.
######## .

#### 2. Slingshot trajectory
```{r, eval=F, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#### Make clusters for Slingshot
## directly pick clusters according to global clusters on fibroblast (resolution=.5)
fib_clu<- fib_clus[,5]
```

```{r, eval=F, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#### slingshot with global clusters of fibroblast.
library(slingshot)
fib_clu_ear <-tapply( names(fib_clu), fib_clu, function(x){ (sum( all_meta[x,1] %in% 'our')*2 + sum( all_meta[x,1] %in% 'asp'))/length(x)   }) # find root by the number of cells from early stage (our*2, asp*1)
fib_lin <- getLineages(fib_umap, fib_clu)
```
```{r, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
#pdf(paste('../etchevers/result/','fib/','gloclu2_glo_slin.pdf',sep='') )
par(cex=2,las=1,mar=rep(0.1,4),lwd=3, pch=16 , xaxt='n',yaxt='n')
plot(fib_umap, col = col6[-1][fib_clu], frame=F, cex=.2)
lines(SlingshotDataSet(fib_lin), lwd = 3, col = 'black', cex= 0.01 )
plot_cen<- sapply( names(fib_clu_ear) , function(x){ c( mean(fib_umap[ names(fib_clu)[fib_clu==x], 1]), mean(fib_umap[ names(fib_clu)[fib_clu==x], 2]) ) })
clu_cex<- fib_clu_ear*3; clu_cex[clu_cex<0.5]<-0.5; clu_cex[clu_cex>2]<-2
for( i in 1:ncol(plot_cen)) points( plot_cen[1,i], plot_cen[2,i] , cex=clu_cex[i] ) # plot node size by percentage of early cells
#dev.off()
```

##### Figure 2.1: slingshot trajectory based on reclustering in the integrated data of fibroblast.

```{r, eval=T, tidy=TRUE, tidy.opts=list(width.cutoff=70)}
# save.image('asp_scrna_blood.rdata.rdata') # save workspace
sessionInfo()
```